QPLIB: A Library of Quadratic Programming Instances

### 2018-08-18 QPLIB-1.0.tex
More typographical items

Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last almost seventy years. At least some of the rationale for this interest is likely due to the fact that QPs are the ``least-nonlinear nonlinear problems''. Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case more easily than would be required to tackle general (Mixed-Integer) Nonlinear Programming ((MI)NLP) problems. This has indeed happened over-and-over again with different algorithmic techniques, such as interior-point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
Similarly, nonconvex continuous QP is perhaps the ``simplest'' class of problems that require features such as spatial enumeration techniques for their solution.
Hence, QPs are both a natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.

In addition, QP, with continuous or integer variables, is arguably a considerably more expressive class than (MI)LP. Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}. Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with quadratic constraints, quadratic objective function, or both. In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree. In turn, every polynomial can be broken down to a system of quadratic expressions. Hence, QP is, in some sense, roughly as expressive as MINLP. This is, in principle, true for MILP as well, but at the cost of much larger and much more complicated formulations. Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost, of MINLP.

@@

The structure of this paper is as follows. In \S\ref{sec:QPbasic} we review the basic notion of QP. In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a new taxonomy of QP that helps us in discussing the (very) different classes of QPs, and \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
Next, \S\ref{sec:lib} describes the process used to obtain the library and its results.
%; the software tools that we have used, and that are freely released together with the library, are discussed in \S\ref{subsec:tools}.
Some conclusions are drawn in \S\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes a simple (QPLIB) file format that encodes all of our examples.

>>>>

Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research for almost seventy years. Part of the rationale for this interest is that QPs are the ``least-nonlinear nonlinear problems''. Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case more easily than would be required to tackle general (Mixed-Integer) Nonlinear Programming ((MI)NLP) problems. This has indeed happened over-and-over again with different algorithmic techniques, such as interior-point methods, active-set methods, e.g., the simplex method, enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
Similarly, nonconvex continuous QP is perhaps the ``simplest'' class of problems that require features such as spatial enumeration techniques for their solution.
Hence, QPs are both a natural basis for developing general techniques for nonconvex NLP, and a very specific class enabling the development of specialized approaches \cite{Dur2010,Burer2012}.

In addition, QP, with continuous or integer variables, is arguably a considerably more expressive class than (MI)LP. Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}. Table \ref{tbl:application} provides a non-exhaustive collection of applications that lead to formulations with quadratic constraints, quadratic objective function, or both. In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree. In turn, every polynomial can be broken down to a system of quadratic expressions. Hence, QP is, in some sense, roughly as expressive as MINLP. This is, in principle, true for MILP as well, but at the cost of much larger and much more complicated formulations. Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost, of MINLP.

@@

The structure of this paper is as follows. In \S\ref{sec:QPbasic} we review the basic notion of QP. In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a new QP taxonomy that helps discuss the (very) different QP classes, and \S\ref{sec:algo} very briefly reviews the QP solution methods and the solvers we have employed.
Next, \S\ref{sec:lib} describes the process used to obtain the library and its results.
%; the software tools that we have used, and that are freely released together with the library, are discussed in \S\ref{subsec:tools}.
Some conclusions are drawn in \S\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes a simple (QPLIB) file format that encodes all of our examples.

### 2018-08-18 QPLIB-2.1.tex
More typographical items

In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function, or some of the constraints, or both, are quadratic functions. More specifically, the problem has the form
%
\begin{align*}
 \min \textrm{or} \max \;\;
 & {\textstyle \half} x^\top Q^0 x + b^0 x + q^0 \\
 \textrm{such that}\;\;
 & c^i_l \leq {\textstyle \half} x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm, \\
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn,  \\
 %
 \textrm{and}\;\;
 & x_j \in \mathbb{Z} & j \in \Mcz,
\end{align*}
%
%\begin{align*}
% \min \;\;
% & x^\top Q^0 x + b^0 x \\
% %
% & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\
% %
% & l_j \leq x_j \leq u_j & j \in \Mcn  \\
% %
% & x_j \in \mathbb{Z} & j \in \Mcz
%\end{align*}
%
where
%
\begin{itemize}
 \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
 %
 \item $x = [x_j]_{j = 1}^n$ %\in \RR^n$
is a finite vector of real variables;
 %
 \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real (Hessian) matrices: since one is only interested in the value of quadratic forms of the type $x^\top Q^i x$, symmetry can be assumed without loss of generality by just replacing off diagonal pairs $Q^i_{hk}$ and $Q^i_{kh}$ with their average $(Q^i_{hk} + Q^i_{kh}) / 2$;
 %
 \item $b^i$, $c_u^i$, $c_l^i$ for $i \in \{ 0 \} \cup \Mcm$, and $q^0$ are, respectively, real $n$-vectors and real constants;
 %
 \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real lower and upper bounds on each variable $x_j$ for $j \in \Mcn$;
 %
 \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$ (i.e., these are the linear constraints, as opposed to the truly quadratic ones); and
 %
 \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
\end{itemize}
%
Due to the presence of integrality requirements on the variables and of quadratic constraints, this class of problems is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP). It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, j \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones. Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the data of the problem, although they may be implied by the other constraints).

The relative flexibility offered by quadratic functions, as opposed e.g.~to linear ones, allows several reformulation techniques to be applicable to this family of problems in order to emphasize different properties of the various components. Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, in particular in the form of binary variables could always be ``hidden'' by introducing (nonconvex) quadratic constraints utilizing the celebrated relationship $x_j \in \{0, 1\} \iff x_j^2 = x_j$. Therefore, when discussing these problems some effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.

>>>>

In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function, or some of the constraints, or both, are quadratic functions. More specifically, the problem has the form
%
\begin{align*}
 \min \textrm{or} \max \;\;
 & {\textstyle \half} x^\top Q^0 x + b^0 x + q^0 \\
 \textrm{such that}\;\;
 & c^i_l \leq {\textstyle \half} x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm, \\
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn,  \\
 %
 \textrm{and}\;\;
 & x_j \in \mathbb{Z} & j \in \Mcz,
\end{align*}
%
%\begin{align*}
% \min \;\;
% & x^\top Q^0 x + b^0 x \\
% %
% & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\
% %
% & l_j \leq x_j \leq u_j & j \in \Mcn  \\
% %
% & x_j \in \mathbb{Z} & j \in \Mcz
%\end{align*}
%
where
%
\begin{itemize}
 \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
 %
 \item $x = [x_j]_{j = 1}^n$ %\in \RR^n$
is a finite vector of real variables;
 %
 \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real (Hessian) matrices: since one is only interested in the value of quadratic forms of the type $x^\top Q^i x$, symmetry can be assumed without loss of generality by just replacing off diagonal pairs $Q^i_{hk}$ and $Q^i_{kh}$ with their average $(Q^i_{hk} + Q^i_{kh}) / 2$;
 %
 \item $b^i$, $c_u^i$, $c_l^i$ for $i \in \{ 0 \} \cup \Mcm$, and $q^0$ are, respectively, real $n$-vectors and real constants;
 %
 \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real lower and upper bounds on each variable $x_j$ for $j \in \Mcn$;
 %
 \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$ (i.e., these are the linear constraints, as opposed to the truly quadratic ones); and
 %
 \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
\end{itemize}
%
Due to the quadratic constraints and the integrality requirements on the variables, this class is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP). It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, j \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous variables. Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the problem data, although bounds may be implied by the other constraints).

The relative flexibility offered by quadratic functions, as opposed, e.g., to linear ones, allows several reformulation techniques to be applicable to this family of problems in order to emphasize different properties of the various components. Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, in particular in the form of binary variables could always be ``hidden'' by introducing (nonconvex) quadratic constraints utilizing the celebrated relationship $x_j \in \{0, 1\} \iff x_j^2 = x_j$. Therefore, when discussing these problems, some effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.

### 2018-08-18 QPLIB-2.2.tex
More typographical items

Despite the apparent simplicity of the definition given in \S\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on fine details of the data. In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties. A taxonomy of QP instances should thus strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods. However, the sheer number of different existing approaches, and the fact that new ones are frequently proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases. This is why, in this paper, we propose an approach that aims at finding a good balance between simplicity and coverage of the main families of computational methods.

@@

The ordering of the values in the previous lists is not irrelevant; in general, problems become ``harder'' when going from left to right. More specifically, for the \textit{O} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, linear objective functions are strictly a special case of diagonal convex quadratic ones (by allowing the diagonal elements all to be zero), the latter are a strict subset of general convex quadratic objectives (by allowing the off-diagonal elements all to be zero), and these are strictly subsets of general nonconvex quadratic ones (since these permit any symmetric Hessian including positive semidefinite ones). The only field for which the containment relationship is not a total order is \textit{V}, for which only the partial orderings
\[
 \mbox{\textit{C}} \subset \mbox{\textit{M}} \subset \mbox{\textit{G}},
 \qquad
 \mbox{\textit{B}} \subset \mbox{\textit{M}} \subset \mbox{\textit{G}},
 \qquad
 \mbox{and}
 \qquad
 \mbox{\textit{B}} \subset \mbox{\textit{I}} \subset \mbox{\textit{G}}
\]
hold. In the following discussion we will repeatedly exploit this by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can be applied as well to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less-general class.

@@

We now give a general discussion about the different problem classes that our proposed taxonomy defines.
For simplicity, we will assume minimization problems for the remaining of this section.
Some problem classes are actually ``too simple'' to make sense in our context. For instance, \textit{D*B} problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
\[
 \textstyle
 \min \; \Big\{ \,
 \sum_{j \in \Mcn} \big( \, {\textstyle \half} Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
 l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcz \, \Big\}
 \;\; .
\]
Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ operations (per variable) by closed-form formul{\ae}, projection and rounding arguments. \emph{A fortiori}, the even simpler cases \textit{L*B}, \textit{D*N} and \textit{L*N} (the latter unbounded unless $b^0 = 0$) will not be discussed here. Similarly, \textit{CCN} are immediately solved by linear algebra techniques, and therefore are of no interest in this context. At the other end of the spectrum, in general QP is a hard problem. Actually, \textit{LIQ}---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e.
\[
 \textstyle
 \min \; \Big\{ \, b^0 x \;:\;
 {\textstyle \half} x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
 \;\; ,
\]
is not only $\mathcal{NP}$-hard, but undecidable \cite{jeroslow}. Hence so are the ``harder'' \{\textit{C},\textit{Q}\}\textit{IQ}.

>>>>

Despite the apparent simplicity of the \S\ref{subsec:notation} definition, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on fine details of the data. In particular, many algorithmic approaches can only be applied to QP when the problem data has specific properties. A taxonomy of QP instances should thus strive to identify a set of properties that an instance should have in order to apply the most relevant computational methods. However, the sheer number of different existing approaches, and the fact that new ones are frequently proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases. This is why, in this paper, we propose an approach that aims at finding a good balance between simplicity and coverage of the main families of computational methods.

@@

The ordering in the preceding lists is relevant; in general, problems become ``harder'' when going from left to right. More specifically, for the \textit{O} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, linear objective functions are strictly a special case of diagonal convex quadratic ones (by allowing the diagonal elements all to be zero), the latter are a strict subset of general convex quadratic objectives (by allowing the off-diagonal elements all to be zero), and these are strictly subsets of general nonconvex quadratic ones (since these permit any symmetric Hessian including positive semidefinite ones). The only field for which the containment relationship is not a total order is \textit{V}, for which only the partial orderings
\[
 \mbox{\textit{C}} \subset \mbox{\textit{M}} \subset \mbox{\textit{G}},
 \qquad
 \mbox{\textit{B}} \subset \mbox{\textit{M}} \subset \mbox{\textit{G}},
 \qquad
 \mbox{and}
 \qquad
 \mbox{\textit{B}} \subset \mbox{\textit{I}} \subset \mbox{\textit{G}}
\]
hold. The following discussion repeatedly exploits this ordering by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can also be applied to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less-general class.

@@

We now give a general discussion about the different problem classes that our proposed taxonomy defines.
For simplicity, this section assumes minimization problems.
Some problem classes are actually ``too simple'' to make sense in our context. For instance, \textit{D*B} problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
\[
 \textstyle
 \min \; \Big\{ \,
 \sum_{j \in \Mcn} \big( \, {\textstyle \half} Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
 l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcz \, \Big\}
 \;\; .
\]
Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ operations (per variable) by closed-form formul{\ae}, projection and rounding arguments. \emph{A fortiori}, the even simpler cases \textit{L*B}, \textit{D*N} and \textit{L*N} (the latter unbounded unless $b^0 = 0$) will not be discussed here. Similarly, \textit{CCN} are immediately solved by linear algebra techniques, and therefore are of no interest in this context. At the other end of the spectrum, in general QP is a hard problem. Actually, \textit{LIQ}---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e.
\[
 \textstyle
 \min \; \Big\{ \, b^0 x \;:\;
 {\textstyle \half} x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
 \;\; ,
\]
is not only $\mathcal{NP}$-hard, but undecidable \cite{jeroslow}. Hence so are the ``harder'' \{\textit{C},\textit{Q}\}\textit{IQ}.

### 2018-08-18 QPLIB-3.tex
More typographical items

In this section we present all the steps we performed in order to build
the new instance library. In \S\ref{subsec:instColl}, we describe the set
of gathered instances, and
in \S\ref{subsec:selection} we present the features used to
classify the instances.
%and discuss issues concerning the format of the instances.
We describe the selection process used to filter the instances, and
graphically present the main features of the selected instances in
\S\ref{subsec:final set}, while in \S\ref{subsec:website}
we provide information on how to access the test collection.

@@

In this section we describe the procedure we adopted to gather the
instances. In January $2014$, we issued an online call for instances
using main international mailing lists of the mathematical
optimization and numerical analysis communities, reaching in this way
a {large set of possibly interested} researchers and practitioners.
The call remained open for ten months, during which we received a large
number of contributions of different nature. The instances we gathered
come both from theoretical studies as well as from real-world
applications.

In addition to these spontaneous contributions we analyzed existing generic
libraries of instances available  on the internet that contain QP
instances. Namely, the libraries from which we gathered instances are
%
\begin{itemize}
 \item the \texttt{BARON} library
 \url{http://www.minlp.com/nlp-and-minlp-test-problems};
%
\item the \texttt{CUTEst} library
 \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest};
%
\item the \texttt{GAMS Performance} libraries
 \url{http://www.gamsworld.org/performance/performlib.htm};
%
\item the \texttt{MacMINLP} library
 \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP};
%
\item the \texttt{Maros-M{\'e}sz{\'a}ros} library
 \url{http://www.doc.ic.ac.uk/~im/00README.QP};
%
\item the \texttt{MINLPLib} library
 \url{http://www.gamsworld.org/minlp/minlplib.htm};
%
\item the \texttt{POLIP} library
 \url{http://polip.zib.de/pipformat.php}.
\end{itemize}

@@

At the end of this process we had gathered more than eight thousand
instances. Three quarters of them contained discrete variables, while
the remainder contained only continuous variables. In more detail,
we gathered $\approx 1800$ Quadratic Binary Linear (\textit{QBL}) instances,
$\approx 2000$ Quadratic Continuous Quadratic (\textit{QCQ}) instances,
and $\approx 2500$ Quadratic General Quadratic (\textit{QGQ}) instances. We
also received $\approx 1000$ Convex General Convex (\textit{CGC}) instances. We
obtained relatively fewer Quadratic Binary Quadratic (\textit{QBQ}), Convex
Continuous Convex (\textit{CCC}) and Convex Mixed Convex (\textit{CMC}) instances,
($\approx 150$, $\approx 200$, and $\approx 200$ instances, respectively).
Finally, we found only $17$ Quadratic Mixed
Linear (\textit{QML}) instances. In the call for instances, no specific format
requirements were imposed for the submissions.

@@

{We now analyze the features of the instances selected to be part of the
 library. In Table \ref{tab:DDDDD}, we provide a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

>>>>

This section presents all the steps we performed to build
the new instance library. In \S\ref{subsec:instColl}, we describe the set
of gathered instances, and
in \S\ref{subsec:selection} we present the features used to
classify the instances.
%and discuss issues concerning the format of the instances.
We describe the selection process used to filter the instances, and
graphically present the main features of the selected instances in
\S\ref{subsec:final set}, while in \S\ref{subsec:website}
we provide information on how to access the test collection.

@@

This section describes the procedure we adopted to gather the
instances. In January $2014$, we issued an online call for instances
using main international mailing lists of the mathematical
optimization and numerical analysis communities, reaching in this way
a {large set of possibly interested} researchers and practitioners.
The call remained open for ten months, during which we received a large
number of contributions of different nature. The instances we gathered
come both from theoretical studies as well as from real-world
applications.

In addition to these spontaneous contributions, we analyzed existing generic
instance libraries available  containing QP
instances. The libraries from which we gathered instances are
%
\begin{itemize}
 \item the \texttt{BARON} library
 \url{http://www.minlp.com/nlp-and-minlp-test-problems};
%
\item the \texttt{CUTEst} library
 \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest};
%
\item the \texttt{GAMS Performance} libraries
 \url{http://www.gamsworld.org/performance/performlib.htm};
%
\item the \texttt{MacMINLP} library
 \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP};
%
\item the \texttt{Maros-M{\'e}sz{\'a}ros} library
 \url{http://www.doc.ic.ac.uk/~im/00README.QP};
%
\item the \texttt{MINLPLib} library
 \url{http://www.gamsworld.org/minlp/minlplib.htm};
%
\item the \texttt{POLIP} library
 \url{http://polip.zib.de/pipformat.php}.
\end{itemize}

@@

At the end of this process, we had gathered more than eight thousand
instances. Three quarters of them contained discrete variables, while
the remainder contained only continuous variables. In more detail,
we gathered $\approx 1800$ Quadratic Binary Linear (\textit{QBL}) instances,
$\approx 2000$ Quadratic Continuous Quadratic (\textit{QCQ}) instances,
and $\approx 2500$ Quadratic General Quadratic (\textit{QGQ}) instances. We
also received $\approx 1000$ Convex General Convex (\textit{CGC}) instances. We
obtained relatively fewer Quadratic Binary Quadratic (\textit{QBQ}), Convex
Continuous Convex (\textit{CCC}) and Convex Mixed Convex (\textit{CMC}) instances,
($\approx 150$, $\approx 200$, and $\approx 200$ instances, respectively).
Finally, we found only $17$ Quadratic Mixed
Linear (\textit{QML}) instances. In the call for instances, no specific format
requirements were imposed for the submissions.

@@

{We now analyze the features of the instances selected to be part of the
 library. Table \ref{tab:DDDDD} provides a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

### 2018-08-18 methods.tex
Update methods.tex
More typographical items

In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (see \S\ref{subsec:solver}). For each approach we briefly describe  the formulation they address according to the classification set out in \S\ref{sec:classification}. We remark that many solvers implement more than one algorithm, which the user can choose at runtime. Moreover, algorithms are typically implemented in different ways within different solvers, so that the same conceptual algorithm can sometimes yield different results or performance measures on the same instances.

@@

The solvers in Table \ref{t:solvers} which implement incomplete methods for NLPs (a problem class containing \textit{QCQ}) are \conopt, \ipopt, \minos, \snopt, and \knitro. Note that all these solvers tackle the more general class of NLP, while we use them only for the considerably more restricted class of QP. Aside from solvers provided by GAMS, there are a number of other, specialized, incomplete QP solvers, such as
\cqp \cite{GoulOrbaRobi13:mpc},
\dqp \cite{GoulRobi16:coap} and
\ooqp \cite{GertWrig03:toms}
for convex problems, and
\bqpd \cite{Flet:2000},
\qpa \cite{GoulToin02i} and
\qpb \cite{ConnGoulOrbaToin:2000},
\qpc \cite{GoulOrbaToin03:toms},
\sqic \cite{Gill2015}
for nonconvex ones.

@@

Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in \textit{QGQ}$\smallsetminus$\textit{CCC}) are based on Branch-and-Bound (BB) \cite{land_doig}. When the BB algorithm is allowed to branch on coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB) \cite{dakin,BeLeLiMaWa08}. BB algorithms require exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice. They can also be used heuristically (forsaking their completeness guarantee) in a number of ways, e.g.~by terminating them early. The following solvers from Table \ref{t:solvers} implement complete BB algorithms for \textit{QGQ} or some subclasses: 
%
\begin{itemize}
 \item \antigone, \baron, \couenne, \lindo, and \scip for \textit{QGQ};
 %
 \item \cplex for \textit{QGL} and \textit{CGC};
%
 \item \textcolor{red}{\gurobi and \xpress for \textit{QBC};}
%
 \item \bonmin, \gurobi, \knitro, \mosek, \sbb, and \xpress for \textit{CGC}.
\end{itemize}
%
We remark that {the solvers \bonmin, \knitro, and \sbb from} the latter category can be used as incomplete solvers for \textit{QGQ}. We also note that \lgo implements an incomplete BB algorithm for \textit{QCQ} by using bounds obtained from sampling.

>>>>

This section provides a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the specific solvers considered in this paper (see \S\ref{subsec:solver}). For each approach, we briefly describe the formulation they address according to the \S\ref{sec:classification} classification. Many solvers implement more than one algorithm, which the user can choose at runtime. Moreover, algorithms are typically implemented in different ways within different solvers, so that the same conceptual algorithm can sometimes yield different results or performance measures on the same instances.

@@

The Table \ref{t:solvers} solvers which implement incomplete methods for NLPs (a problem class containing \textit{QCQ}) are \conopt, \ipopt, \minos, \snopt, and \knitro. Note that all these solvers tackle the more general class of NLP, while we use them only for the considerably more restricted QP class. Aside from solvers provided by GAMS, there are a number of other, specialized, incomplete QP solvers, such as
\cqp \cite{GoulOrbaRobi13:mpc},
\dqp \cite{GoulRobi16:coap} and
\ooqp \cite{GertWrig03:toms}
for convex problems, and
\bqpd \cite{Flet:2000},
\qpa \cite{GoulToin02i} and
\qpb \cite{ConnGoulOrbaToin:2000},
\qpc \cite{GoulOrbaToin03:toms},
\sqic \cite{Gill2015}
for nonconvex ones.

@@

Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in \textit{QGQ}$\smallsetminus$\textit{CCC}) are based on Branch-and-Bound (BB) \cite{land_doig}. When the BB algorithm is allowed to branch on coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB) \cite{dakin,BeLeLiMaWa08}. BB algorithms require exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice. They can also be used heuristically (forsaking their completeness guarantee) in a number of ways, e.g.~by terminating them early. The following solvers from Table \ref{t:solvers} implement complete BB algorithms for \textit{QGQ} or some subclasses: 
%
\begin{itemize}
 \item \antigone, \baron, \couenne, \lindo, and \scip for \textit{QGQ};
 %
 \item \cplex for \textit{QGL} and \textit{CGC};
%
 \item {\gurobi and \xpress for \textit{QBC};}
%
 \item \bonmin, \gurobi, \knitro, \mosek, \sbb, and \xpress for \textit{CGC}.
\end{itemize}
%
We remark that {the solvers \bonmin, \knitro, and \sbb from} the latter category can be used as incomplete solvers for \textit{QGQ}. We also note that \lgo implements an incomplete BB algorithm for \textit{QCQ} by using bounds obtained from sampling.
